---
title: "HW3"
output: html_document
---

Load data, ggplot2:
```{r}
data(tips, package="reshape2")
library("ggplot2")
```

##1.
```{r}
numParty <- nrow(tips)
numParty
```

It appears that there were 244 parties served by the waiter.

##2.
```{r}
str(tips)
```

You can see here that there are numbers, strings, and ints in this data frame.

##3.
```{r}
tips$day <- factor(tips$day, levels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))
day <- tips[order(tips$day), "day"]
day
```

##4.
```{r}
female <- tips[(tips$sex == 'Female'),]
female5 <- female[(female$size >= 5),]
female5
```

##5.
```{r}
ggplot(tips, aes(x=tip, y=total_bill)) + geom_point() + stat_ellipse(type="norm") + ggtitle("Tip Amount vs. Total Bill") + theme(plot.title=element_text(hjust=0.5))
```

I included an ellipse around the normal data to show help explain anomalies. In general, the relationship here is that, as the total bill increases, tip rate also increases. Two anomalies I see in this graph would be some of the points that are closer to 0 on the x-axis, but higher on the y-axis, near the $30-40 range. These are very low tips, especially for the trend shown in this scatterplot. Another anomaly would be the point that is at approximately (5.2, 7.5) here. This tip is of a very high percentage of the total bill, which makes it out of the ordinary here.

##6.
```{r}
tips$tiprate <- tips$tip / tips$total_bill
head(tips, 10) # to show new column
avgTip <- mean(tips$tiprate)
avgTip # average tip rate is approximately 16%
```

##7.
```{r}
tips$time <- factor(tips$time, levels = c("Lunch", "Dinner"))
temp <- tips[order(tips$time),] # to hold new ordering
p <- ggplot(temp, aes(x=time, y=tiprate)) + geom_boxplot(outlier.color="red", outlier.shape=8, outlier.size=2)
p <- p + ggtitle("Time of Day vs. Tip Rate") + theme(plot.title=element_text(hjust=0.5))
p
```

##8
The relation between tip and total bill here is easily represented by the tiprate column, however I will use both for the scatterplots. From here I will try and find meaning between smoking and gender of person who pays to the tip rate. I used the ggpubr library, which I found useful during an internship I had this past summer. It is helpful for arranging graphs.
```{r}
library("ggpubr")

# I had to make some dataframes so these scatterplots could be created easier

female <- tips[(tips$sex == 'Female'),]
f_noSmoke <- female[(female$smoker == 'No'),]
f_yesSmoke <- female[(female$smoker == 'Yes'),]
male <- tips[(tips$sex == 'Male'),]
m_noSmoke <- male[(male$smoker == "No"),]
m_yesSmoke <- male[(male$smoker == "Yes"),]

# Now making scatterplots to then turn into a grid

f_no <- ggplot(f_noSmoke, aes(x=tip, y=total_bill)) + geom_point() + geom_smooth()
f_yes <- ggplot(f_yesSmoke, aes(x=tip, y=total_bill)) + geom_point() + geom_smooth()
m_no <- ggplot(m_noSmoke, aes(x=tip, y=total_bill)) + geom_point() + geom_smooth()
m_yes <- ggplot(m_yesSmoke, aes(x=tip, y=total_bill)) + geom_point() + geom_smooth()

# Using ggarrange to organize into grid

plots <- ggarrange(m_no, m_yes, f_no, f_yes, labels=c("Male Non-Smoker", "Male Smoker", "Female Non-Smoker", "Female Smoker"), font.label=list(size = 10, color="black"), ncol=2,nrow=2)
plots
```

From this result, here are my interpretations:
I think that females, when paying, tend to have lower overall bills and therefore lower tips. However, it seems that they have a lower tip rate. A flatter regression line will show that the tip rate is higher, and both male smokers and non-smokers exhibit a flatter lines than their female counterparts, which implies a higher tip rate. When looking at smoking specifically, it appears that the smokers tend to tip better than non-smokers, based on the shape of the regression lines. Overall, there is some interesting information to draw from these graphs, which is easy to compare when they are side-by-side.

##9.
I decided to use a bar plot to describe and analyze this data.
```{r}
# First creating a vector of numerical values so the bar plot has some meaning (can reuse male, female dataframes from earlier)

fT <- female[(female$day == "Thur"),]
fF <- female[(female$day == "Fri"),]
fSa <- female[(female$day == "Sat"),]
fSu <- female[(female$day == "Sun"),]
mT <- male[(male$day == "Thurs"),]
mF <- male[(male$day == "Fri"),]
mSa <- male[(male$day == "Sat"),]
mSu <- male[(male$day == "Sun"),]

yvals <- c(nrow(fT),nrow(mT),nrow(fF),nrow(mF),nrow(fSa),nrow(mSa),nrow(fSu),nrow(mSu))
xvals <- c("female_Thurs","male_Thurs","female_Fri","male_Fri","female_Sat","male_Sat","female_Sun","male_Sun")
toPlot <- data.frame("Days" = xvals, "Number" = yvals)
positions <- c("female_Thurs","male_Thurs","female_Fri","male_Fri","female_Sat","male_Sat","female_Sun","male_Sun")
ggplot(toPlot, aes(Days, Number)) + geom_bar(stat="identity") + scale_x_discrete(limits=positions)
```

From this graph, I can see that the percentage of females paying is much higher during the week, but males paying over the weekend is drastically higher.